Analysis and Figures for Gallego et al, 2020

### 2020-09-27 Manuscript/Response.to.reviewers.tex

\textbf{ 
 Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters}

\textbf {Response to reviewers}

\vspace*{1em}

We are thankful for the feedback, comments and suggestions on our manuscript RSPB-2019-2901  "Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters". Find below our response to their suggestions, and how we implement them in our manuscript.

@@

Figure 2 should be exactly the same as the first part of the Supplemental figure - A thorough review of the code that generates the supplemental figures revealed a small difference in the input dataset. We thank the reviewer for noticing this, and now the code that generates each of the figures is consistent.

\item{As Figure 3 is divided into sections A, B, and C, these ought to be referred to in the text (see below). The text is very small and hard to read in the review version I have.}

\textbf{Response:}

In an effort to declutter Figure 3, and following suggestions from reviewer 1, we have focused the figure in those species showing a more dramatic change in suitability. The new version of the figure has three panels (A, B, C; one per phylum) displaying three aspects of the data (.1, .2 and .3) and we refer to them specifically in the text.

@@

Done.

\item{Line 99: The value of predicting which cluster (?community?) a given set of environmental parameters will fall within is not entirely clear to me. Is there a consistent set of taxa that make of up these clusters across time and space? Can specific groups of taxa be deduced from the ?community? number? Do the environmental variables accurately predict which set of taxa will be present? This section needs a bit more clarity.}

\textbf{Response:}

In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions. However, it is not that taxa occur exclusively in different clusters;  differences in relative proportions of species will also contribute to cluster dissimilarities. We see the clusters as different communities, in the sense that they are characteristic \textit{mixes} of taxa that occur under different conditions.

@@

Done.

\item{Line 121-133: When referring to specific taxa as winners or losers, I was expecting to see them in Figure 3B, but not all of them were there. Or if it did appear, it was only at one site, but it was mentioned as being an overall winner or loser.}

\textbf{Response:}

 We thank the reviewer for clarifying this point. We have now decluttered Figure 3; and highlight taxa that are now mentioned in the text: the Dinoflagellates \textit{Karlodinium}, Gymnodinaceae sp. and \textit{Alexandrium}; the Diatoms \textit{Stephanodiscus}, \textit{Nitzschia}, \textit{Coscinodiscus} and \textit{Chaetoceros}; and the Haptophytes \textit{Emiliania huxlei} and \textit{Chrysochromulina} sp.

@@

Our effort has been on visualizing species with more pronounced change, not to highlight species that respond in contrasting ways in the San Juan Islands and the Hood Canal. Most taxa react in the same direction in both regions; with a few exceptions in which species are almost absent on one study area.

\item{Line 123: Coscinodiscus and Ditylum seem to refer to Hood Canal. This ought to specified as before.}

\textbf{Response:}

\textit{Coscinodiscus} sees its probability of presence reduced in both Areas. The same happens for \textit{D. brightwelli}, although our approach for a more readable figure has seen this taxon drop into the Supplementary figures.

@@

 A reworded version of this paragraph is now at the end of the discussion section.

\item{Line 147-148: Need a citation for Coscinodiscus as key fish food.}

\textbf{Response:}

 Agreed: We now cite the work of J. Zamon (2002) and S. Pasquaud and colleagues (2010) to support the role of \textit{Coscinodiscus} in the food chain.

\item{Line 178: Spell out harmful algal bloom the first time it's used.}

@@

 Agreed - it will definitely have an effect on the `fate` of the photosynthesized carbon- a quickly sunk carbon will join marine sediments instead of being recycled in the upper layers of the water masses. We now cite the work of Polimene and colleagues on the role of phytoplankton and carbon sequestration.

\item{Line 204-205: Citation needed for fish mortality. Eutrophication seems like a cause of blooms, not a result of them. What are the potential cascade effects on benthic-pelagic coupling? The term here on its own isn?t really saying anything.}

\textbf{Response:}

 In their work, D. Anderson and colleagues cite a range of consequences of increased HAB blooms, including fish mortality and anoxia - and not eutrophication as we wrongly pointed out. Thanks for noticing. On the second argument - we saw potential consequences in two ways: by affecting recruiting of benthic species with pelagic larvae with reduced diatom diversity/biomass; and by changing the nature of sinking matter for filter feeders. We realize that bentho-pelagic coupling is not the only source of carbon for benthic communities in the area of study so we have decided to remove this assertion from the discussion.

@@

Done.

\item{Line 267-268: Which index of abundance from Kelly et al. (2019) was used? That paper presents several and though it declares one to be more useful, I think it would be helpful in this manuscript to state which one was used.}

\textbf{Response:}

We used the eDNAindex (the proportion of reads assigned to each taxa per sample, scaled to the maximum proportion that taxa achieved). We have now specified this and provide the R implementation of this index in the Supplementary material(line 128).

@@

\item{Line 273-282: It's hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area. I think this section could use a bit of clarification, and all the code and raw data should be given (the provided github link is broken).}

\textbf{Response:}

We apologize for the broken github code - We have clarified how we simulated the future projection in the Supplementary Information - Also see response to reviewer 1 on line 122 - 137 of this letter.

### 2020-11-04 Manuscript/Response.to.reviewers.tex

\item{3.      The manuscript suffers from a lack of transparent results pertaining to the sequencing and bioinformatic methods. How many samples failed to pass QC? How many reads per sample were the analyses based on? How many samples did you collect from each site? I think this manuscript will benefit from including a table in the main text that summarizes some of these data.}

\textbf{Response:}

We had intended to strike a balance between readability (on one hand) and details that are important to specialists (on the other). Plainly, the first version of the manuscript erred too far on the side of brevity, and we have now included many such details in the main text.

Moreover, due to a technical error, the raw data and scripts were not available to reviewers in the first submission -- our sincere apologies. All data and scripts are now available in an open github repository \url{https://github.com/ramongallego/eDNA.and.Ocean.Acidification.Gallego.et.al.2020} (raw fastq files are in a google drive \href{https://drive.google.com/drive/folders/1XtuDMdlnk9V2acNiPJo5T6O1Tp0KjaCz?usp=sharing}{link}, because of their size). We have strived for transparency and repeatability throughout.

@@

In fact, this reviewer's concern -- which another reviewer shared -- led us to simplify our analysis in general, dropping a seasonality term from our model. This helped clarify the core findings, and we are grateful for the feedback.

\end{itemize}

\vspace*{1em}
\textbf{\large{ Reviewer 2}}:

\begin{itemize}
\item{Figures: the colors red and blue are used both to distinguish time periods (2019/2095) and locality (San Juan/Hood Canal) in different figures. Figure 2 legend doesn't say which color is which (presumably they are the same as in Figure 1). Use consistent colors for categories across figures and use different colors for time period and locality, if possible.}

### 2020-11-04 Manuscript/What.lies.ahead.reviewed.tex
Final revision

\subsection{Background}

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification. A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}.

@@

\section{Methods}
\subsection*{Sampling}

@@

@@

\subsection*{Results}
\underline{Variation in Carbonate Chemistry and in Ecological Communities}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supplementary Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

\subsection*{Discussion}

@@

 Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}. Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling and carbon sequestration \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking, polimene2017biological}. Although the north Atlantic has shown an increase in diatom abundance \cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}. Furthermore, locally focused models support an increase in dinoflagellate dominance with climate change, particularly during summer months \cite{khangaonkar2019salish}.

@@

\section{Acknowledgments}

>>>>

\section*{Background}

@@

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification. A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2020environmental}.

@@

\section*{Methods}
\subsection*{Sampling}

@@

All raw data used in this manuscript and the scripts that generate the final figures and analysis can be found at https://github.com/ramongallego/eDNA.and.Ocean.Acidification.Gallego.et.al.2020

@@

\section*{Results}
\underline{Variation in Carbonate Chemistry and in Ecological Communities}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supplementary Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

\section*{Discussion}

@@

 Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}. Recent studies using mesocosm manipulations also suggest a foundational shift in community composition in that direction \cite{Hyun_2020, Crummett_2020}. Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling and carbon sequestration \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking, polimene2017biological}. Although the north Atlantic has shown an increase in diatom abundance \cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}. Furthermore, locally focused models support an increase in dinoflagellate dominance with climate change, particularly during summer months \cite{khangaonkar2019salish}.

@@

\section*{Acknowledgments}

### 2020-11-04 Scripts/Manuscript.Main.Analysis.Rmd
Final revision

Fig1
ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 12, dpi = "retina")

@@

ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina")

@@

(left|right)+plot_layout(widths = c(2,2))  -> plot 
   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 14)

>>>>

Fig1
ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 12, dpi = "retina")
ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.pdf"), width = 14, height = 12, dpi = "retina")

@@

ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina") 
ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".tiff")),
         width = 14,height = 11, dpi = "retina")

@@

(left|right)+plot_layout(widths = c(2,2))  -> plot 
   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 14)
   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.tiff")), units = "in", dpi = "retina", width = 20, height = 14)
 ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.pdf")), units = "in", dpi = "retina", width = 20, height = 14)
```
